% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
% EDIT THIS
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 

\department{Department of Applied Mathematics}
\title{Statistical modeling of COVID-19 time series}
\authorGN{Oleh} %author's given name/names
\authorFN{Kuznetsov} %author's surname
\author{Kuznetsov Oleh} %author's name without academic degrees
\authorWithDegrees{Kuznetsov Oleh} %author's name with academic degrees
\supervisor{Ing. Dedecius Kamil, Ph.D.}
\acknowledgements{I express my sincere gratitude to my supervisor for this thesis Ing. Dedecius Kamil, Ph.D. It was thanks to him that I became interested in statistics and was motivated to study the topics covered in this work. The advice, guidance, and information received from him helped me deepen into this problem and achieve good results.

I am deeply grateful to my family (and especially my dear mother, father, and sister) for their support and opportunity to move to the Czech Republic and study at the Czech Technical University in Prague.

Last but not least, I would like to frankly acknowledge my friends for being with me during my studies.
}
\abstractCS{V letech 2020-2021 zažívá celý svět potíže způsobené globální epidemií nového viru SARS-CoV-2. Cílem této práce je statistická analýza časových řad souvisejících s pandemií COVID-19 v České republice a zkoumání využitelnosti modelů Facebook Prophet a Seasonal Autoregressive Integrated Moving Average (SARIMA) k analýze a predikci vývoje pandemických procesů. Tato práce propojuje globální změny, které nastaly ve vybraných časových řadách, s poskytnutými vládními omezeními (lockdown, nošení obličejové masky apod.).}
\abstractEN{In the years 2020-2021, the whole world is experiencing difficulties caused by the global epidemic of the new SARS-CoV-2 virus. This thesis aims at the statistical analysis of the time series related to the COVID-19 pandemic in the Czech Republic. It explores the usability of the Facebook Prophet and Seasonal Autoregressive Integrated Moving Average (SARIMA) models to analyze and predict the development of pandemic processes. Moreover, this thesis connects the global changes that occurred in the selected time series with the provided government restrictions (lockdown, wearing a face mask, and so on).}
\placeForDeclarationOfAuthenticity{Prague}
\keywordsCS{COVID-19, časové řady, prognózy časových řad, analýza časových řad, statistické modelování časových řad, Facebook Prophet, SARIMA  \clearpage}
\keywordsEN{COVID-19, Time series, Time series forecasting, Time series analysis, Statistical time series modeling, Facebook Prophet, SARIMA}
\declarationOfAuthenticityOption{1} %select as appropriate, according to the desired license (integer 1-6)
% \website{http://site.example/thesis} %optional thesis URL
